Very simplified likelihood

### 2024-09-17 inst/very_simplified_likelihood.Rmd
bugfix

```{r definelikeinsilico}
likeinsilico <- function(vec,M){
  alpha <- vec[1]
  beta  <- vec[2]
  S <- 0
  for(i in seq_len(nrow(M))){
    x <- M[i,1]
    LO <- alpha + beta * x
    p <- exp(LO)/(1+exp(LO))
    S <- S + f_vec_arn(p,R=M[i,2],n=M[i,3]-1,log=TRUE)
  }
return(S)
}
```

@@

```{r doZcontour}
Z <- Z-max(Z)
Z <- pmax(Z,-10)
contour(a,b,Z,levels=-c(0:10,15,20,30))
points(alpha_true,beta_true,pch=16,cex=4)
jj <- optim_ab$par
points(jj[1],jj[2],pch=4,cex=4,col='red')
grid()
persp(a,b,Z)
```

@@

```{r plotolymp}
plotolymp <- function(...){
plot(a,L,type='l',ylab="log-likelihood",lwd=2)
abline(h=c(0,-2))
ahat <- optolym$maximum
segments(x0=ahat,y0=-1.5,y1=0)
text(ahat,-0.91,expression(hat(a)==0.573),pos=2)
abline(v=0.5,lty=3)

segments(x0=a_lower,y0=-1.5,y1=-3.5)
segments(x0=a_upper,y0=-1.5,y1=-3.5)
grid()
text(x=a_lower,y=-3,"0.380",pos=4)
text(x=a_upper,y=-3,"0.563",pos=2)
}
plotolymp()
pdf(file="plotolymp.pdf")
plotolymp()
dev.off()
```

@@

```{r calculate_parksupp,cache=TRUE}
S_parkrun <- function(a){f_vec_arn(a,R=results$place,results$runners-1,log=TRUE)}
(optparkrun <- optimize(S_parkrun, c(0.2,0.8),maximum=TRUE))

@@

```{r splitresults}
supp2 <- function(vec,place,runners){
  a_stirling <- vec[1]
  a_auckland <- vec[2]
  out <- 0
  for(i in 1:3){
    out <- out + f_vec_a(a_stirling,r=place[i],runners[i])
  }
  for(i in 4:10){
    out <- out + f_vec_a(a_auckland,r=place[i],runners[i])
  }
  return(out)
}
```

@@

```{r,label=plotbothlikes}
parkrun_contour <- function(...){
    par(pty='s')
    contour(x,x,L,asp=1,xlim=range(x),ylim=range(x),levels=-(0:5),xlab="Auckland",ylab="Stirling")
    abline(0,1)
}
parkrun_contour()
pdf(file="parkruncontour.pdf")
parkrun_contour()
dev.off()
```

@@

```{r focaleducation}
rank <- c( 9, 7, 2, 3,2 ,8 ,5 ,4 ,9 )
class_size <- c(12, 17,23, 9,13,14,13,12,15)
course <- c("rings and modules","group theory","calculus","linear algebra",
"differential equations","topology","special relativity","fluid mechanics","Lie algebra")
category=c("pure","pure","applied","applied","applied","pure","applied","applied","pure")
data.frame(course,category,rank,class_size)
```

@@

>>>>

```{r definelikeinsilico}
likeinsilico <- function(vec,M){
  alpha <- vec[1]
  beta  <- vec[2]
  S <- 0
  for(i in seq_len(nrow(M))){
    x <- M[i,1]
    LO <- alpha + beta * x
    p <- exp(LO)/(1+exp(LO))
    S <- S + f_vec_arn(p,r=M[i,2],n=M[i,3]-1)
  }
return(S)
}
```

@@

```{r doZcontour}
Z <- Z-max(Z)
Z <- pmax(Z,-10)
contour(a,b,Z,levels=-c(0:10,15,20,30))
points(alpha_true,beta_true,pch=16,cex=4)
jj <- optim_ab$par
points(jj[1],jj[2],pch=4,cex=4,col='red')
grid()
#persp(a,b,Z)
```

@@

```{r plotolymp}
plotolymp <- function(...){
plot(a,L,type='l',ylab="log-likelihood",lwd=2)
abline(h=c(0,-2))
ahat <- optolym$maximum
segments(x0=ahat,y0=-1.5,y1=0)
text(ahat,-0.91,expression(hat(a)==0.511),pos=2)
abline(v=0.5,lty=3)

segments(x0=a_lower,y0=-1.5,y1=-3.5)
segments(x0=a_upper,y0=-1.5,y1=-3.5)
grid()
text(x=a_lower,y=-3,"0.414",pos=4)
text(x=a_upper,y=-3,"0.602",pos=2)
}
plotolymp()
pdf(file="plotolymp.pdf")
plotolymp()
dev.off()
```

@@

```{r calculate_parksupp,cache=TRUE}
S_parkrun <- function(a){f_vec_arn(a, r=results$place, results$runners)}
(optparkrun <- optimize(S_parkrun, c(0.2,0.8),maximum=TRUE))
```

@@

```{r splitresults}
supp2 <- function(vec,place,runners){
  a_stirling <- vec[1]
  a_auckland <- vec[2]
  out <- 0
  for(i in 1:12){
    out <- out + f_vec_a(a_stirling,r=place[i],runners[i])
  }
  for(i in 13:21){
    out <- out + f_vec_a(a_auckland,r=place[i],runners[i])
  }
  return(out)
}
```

@@

```{r,label=plotbothlikes}
parkrun_contour <- function(...){
    par(pty='s')
    contour(x,x,L,asp=1,xlim=range(x),ylim=range(x),levels=-(0:5),xlab="2023",ylab="2024")
    abline(0,1)
}
parkrun_contour()
pdf(file="parkruncontour.pdf")
parkrun_contour()
dev.off()
```

@@

```{r focaleducation}
rank <- c( 9, 7, 2, 3,2 ,8 ,5 ,4 ,9 )
class_size <- c(12, 17,23, 9,13,14,13,12,15)
course <- c("rings and modules","group theory","calculus","linear algebra",
"differential equations","topology","special relativity","fluid mechanics","Lie algebra")
category <- c("pure","pure","applied","applied","applied","pure","applied","applied","pure")
data.frame(course,category,rank,class_size)
```

@@

```{r dofishereducation}
pure_betterthanme    <- sum(rank[wp]-1)
pure_worsethanme     <- sum(class_size[wp] - rank[wp])
applied_betterthanme <- sum(rank[wa]-1)
applied_worsethanme  <-  sum(class_size[wa] - rank[wa])
M <- matrix(c(pure_betterthanme,pure_worsethanme,applied_betterthanme,applied_worsethanme),2,2,byrow=TRUE)
dimnames(M) <- list(type=c("pure","applied"),beatsme=c(T,F))
M
fisher.test(M)
```
